---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading library
```{r}
library(data.table)
library(dplyr,warn.conflicts = FALSE)
library(ggplot2)
library(lubridate,warn.conflicts = FALSE)
library(magrittr)
library(ggpubr)
```


## Loading and preprocessing the data
```{r}
dt<-fread(unzip("activity.zip"))
dt[,date:=ymd(date)]

```
## What is mean total number of steps taken per day?
```{r}
dtsum<-dt[,.(mean=mean(steps,na.rm = TRUE),median=as.double(median(steps,na.rm = TRUE)),total=sum(steps,na.rm = TRUE),maxinterval=interval[which.max(steps)]),by=date]
dtsum
qplot(x=total,data=dtsum,geom="histogram")
```

mean and median of total number of steps taken per day are:
```{r}
dtsum[,c(mean(total),median(total))]
```

## What is the average daily activity pattern?
```{r}
dtsumbyinterval<-dt[,.(mean=mean(steps,na.rm = TRUE)),by=interval]
qplot(x=interval,y=mean,data=dtsumbyinterval,geom="line")

```

## What is the maximum number of steps interval?
```{r}
dtsumbyinterval[which.max(mean),interval]
```

## Imputing missing values
Total number of NAs
```{r}
summary(dt)
```

Missing value pattern
```{r}
dt[,.(N=.N,noNA=sum(is.na(steps))),by=date] 

```
 We can know NA is due to no record on particular day rather than random NA. So it is better to fill NA with mean of the 5 minutes interval
```{r}
dtimp<-dt
NAintervallist<-dtimp[is.na(steps),interval]
fillvalue<-sapply(NAintervallist,function(x){
  return(dtsumbyinterval[interval==x,mean])
})
class(dtimp$steps)<-"double"
dtimp[is.na(steps),steps:=fillvalue]

dtimpsum<-dtimp[,.(mean=mean(steps),median=as.double(median(steps,na.rm = TRUE))
                   ,total=sum(steps),maxinterval=interval[which.max(steps)]),by=date]
dtimpsumbyinterval<-dtimp[,.(mean=mean(steps)),by=interval]
```

After imputing all NAs, the histogram of total steps per day is as below:
```{r}
qplot(x=total,data=dtimpsum,geom="histogram")
print("mean and median of total number of steps taken per day are")
dtimpsum[,c(mean(total),median(total))]
```

The mean and median difference before and after imputing NAs are:
```{r}
dif<-dtimpsum[,c(mean(total),median(total))]-dtsum[,c(mean(total),median(total))]
dif
if(dif>0){
  print("The result estimated total number of steps per day will increase")
  
}
if(dif<0){
  print("The result estimated total number of steps per day will decrease")
}
if(dif==0){
  print("The result estimated total number of steps has no change")
}

```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
dtimp[(wday(date)==1)|(wday(date)==7),weekend:="weekend"]
dtimp[!(wday(date)==1)|(wday(date)==7),weekend:="weekday"]
dtimpsumbyinterval<-dtimp[,.(mean=mean(steps)),by=c("interval","weekend")]
qplot(x=interval,y=mean,data=dtimpsumbyinterval,facets = (weekend~.),geom="line")
```


